\section{Introduction}

Over the past two decades, the field of \emph{network algorithmics}~\cite{Varghese04Algorithmics}
has emerged as a rich area of research. The types of problems that network
algorithmics address include, but are not limited to, packet classification,
queue management, traffic shaping, switching and routing, network measurement,
and data streaming analysis. Many of such solutions have been successfully
deployed in commercial networks, including solutions for Internet routers,
security apparatus, and measurement devices. 

In general, network operators
would like solutions that are robust under a wide variety of, \emph{often
unforeseen}, operating conditions. However, the performance of a network appliance
design or the accuracy of an estimation method is often dependent on workload
uncertainties that are beyond the control of the operator. Unfortunately,
applicable mathematics for the rigorous analysis of the worst-case stochastic behaviors of
network algorithmics solutions under arbitrary workloads is largely lacking.

Understanding how a solution would behave in the worst-case, not just in
the typical case, is important for two reasons. First, with suitable mathematics
to characterize worst-case workloads, we can design solutions that will work
well under any operating conditions, including those in which an adversary is
trying to break the system, or under unexpected changes in the usage pattern.
Second, more often than not, we have, surprisingly, found in our past efforts
that delivering solutions that can guarantee high performance under the worst-case
conditions cost only slightly more than designs that don't. However, coming up
with such solutions hinges on our ability to understand the characteristics
of the worst-case scenarios so that we can design around them.

Large deviation theory~\cite{Dembo98LDV} on $\Real$ is concerned with the probability that
the sum of some random variables $S := \sum_{i=1}^n X_i$ will exceed a given
threshold $x$, which in our contexts may correspond to processing or network capacity,
resource constraint, or tolerable error bound. In worst-case large deviation problems, 
these random variables $X_1, X_2, \cdots, X_n$
are the functions of some large parameter vector $\bq := \lbrack q_1, q_2, \cdots, q_m\rbrack$,
and we wish to find the worst-case probability tail bounds of
$S(\bq) := \sum_{i=1}^n X_i(\bq)$ under all possible parameter
settings $\bq \in Q$. In other words, we would like to compute
$\max_{\bq\in Q} \Pr[S(\bq) \ge x]$.

Establishing such worst-case tail bounds is important not only for network
security applications where an adversary is in full or partial control of
this vector, but also for non-security applications where we would like to
know the worst-case system performance under all operating conditions or workloads.
Obtaining such worst-case bounds is very difficult because the parameter
space $Q$ is typically gigantic. Although establishing tail bounds is
often straightforward through Chernoff bounding
techniques when a particular parameter setting $\bq$ is given,
such bounds typically cannot be expressed as a closed form function of $\bq$,
rendering conventional optimization techniques powerless for maximizing $\Pr[S(\bq) \ge x]$.
Without worst-case large deviation machineries that have been or are being developed, the only
conceivable option would be to enumerate all parameter settings $\bq\in Q$, but repeating tail
bound analysis over the entire parameter space $Q$ is usually computationally prohibitive.

Whether determining the worst case input or bounding the
behavior of randomized algorithms, \textit{stochastic orders} can help in understanding the worst-case behavior of proposed
solutions in network algorithmics. The motivation of this paper is to provide an overview to researchers in
network algorithmics on stochastic orders.

A stochastic order basically defines when a random variable is ``larger'' than another. The precise definition
of ``larger'' than will depend on the order. 
There are many types of stochastic orders that can be defined on random variables. Two chief references are
M\"{u}ller and Stoyan \cite{Muller02Risk}, and Shaked and Shantikumar \cite{Shaked07Sorders}. Of importance
to us are majorization and Schur convexity, supermodular and convex orderings. We will also briefly mention about negatively associated
random variables and how they relate to network algorithmics.

In this paper, we will summarize some useful classifications of distributions and stochastic orders to aid
in the design of randomized algorithms in network algorithmics. 
In particular, we cover:
\begin{enumerate}
\item majorization and Schur convexity, 
\item negative association,
\item convex ordering and supermodularity, and
\item log-concave distributions.
\end{enumerate}
We show how these can be applied to various applications in network algorithmics. Most of the 
results shown here can be found in literature, and so we endeavor to reference the original 
paper where the result originated from. 





%High-speed networks pose a challenge to performing
%operations, such as measurement, on network traffic, due to three main
%factors (also encountered in Big Data applications): volume, velocity and variety. Volume
%and velocity are self-evident: it is common for today's backbone networks to run around 
%100 Gbps, and this speed is likely to be surpassed in the near future. Variety appears in 
%the different forms of traffic an operator would like to measure, for instance, traffic
%flows of particular applications such as Voice-over-IP (VoIP), or to flag anomalies
%in traffic.
%
%On a 100 Gbps link, several Terabytes of data are being transmitted per hour.
%Clearly, the \naive solution of performing operations by first storing all traffic is prohibitively expensive, both
%from a computational and financial standpoint. Network algorithmics, coined by 
%Varghese \cite{Varghese04Algorithmics}, is a field that arose to address these challenges. 
%Some problems network algorithmics has addressed include, but not limited to, packet classification
%(such as deep packet inspection), queue management, traffic shaping, switching and routing, network
%measurement, and data streaming analysis. The solutions from network algorithmics have been 
%successfully deployed in commercial networks, such as new switching architectures, security apparatus, and
%measurement devices. As network speeds increase, couple with increasing traffic demand, network
%algorithmics will be even more prominent in the near future.
%
%As the inputs, \ie traffic streams, into these solutions are 
%often arbitrary, an important issue in network algorithmics is to deal with the worst-case scenario the solutions
%could face. Even though the solution itself has a deterministic behavior, how it behaves will be largely
%determined by the stochastic behavior of the inputs. For instance, in switching, a scheduling policy with 
%deterministic behavior is encouraged because deterministic behavior simplifies the design of the 
%on-chip controller that implements the scheduling policy. The simple yet effective deterministic 
%round-robin policy (with priority queues) of the well-known iSLIP \cite{McKeown99iSlip} 
%scheduling policy explains its success in widespread implementation on Cisco's routers. However,
%analysis of its worst-case behavior is determined by its interaction with specific traffic workloads.
%
%Moreover, many solutions in network algorithmics rely on \textit{randomness} itself. 
%The dual requirements of dealing with high speeds constrained by a small amount of memory necessitates
%the use of compact summaries with \textit{randomized} logic in network measurement. 
%As the name suggests, these randomized algorithms 
%utilizes randomness in its operations. One famous example for testing the membership of an item in a set is the Bloom filter
%\cite{Bloom70Filter}. The Bloom filter enables the compact representation of a set for querying. 
%Suppose a set of items are stored in the filter. If an item is in the set of items, querying for 
%the item will always return a ``Yes''. However, if we have a query for an item \textbf{not} found in the
%set, there is a small probability that the answer will be ``Yes''. This error is known as a false positive.  
%In this sense, the Bloom filter trades off some storage space with a small probability of \textit{error}. 
%
%Essentially, this is the crux of (Monte Carlo) randomized algorithms: ensure a bounded resource (for example,
%space), but trade this off with some error on its operations \cite{MitzenmacherProb05}. 
%Randomized algorithms in network algorithmics require probability bounds on these errors (or ``bad'' events) occurring. 
%
%In both cases above, \textit{stochastic orders} can help in understanding the worst-case behavior of proposed
%solutions in network algorithmics. The motivation of this paper is to provide an overview to researchers in
%network algorithmics on stochastic orders.
%
%A stochastic order basically defines when a random variable is ``larger'' than another. The precise definition
%of ``larger'' than will depend on the order. 
%There are many types of stochastic orders that can be defined on random variables. Two chief references are
%M\"{u}ller and Stoyan \cite{Muller02Risk}, and Shaked and Shantikumar \cite{Shaked07Sorders}. Of importance
%to us are supermodular and convex orderings. We will also briefly mention about negatively associated
%random variables and how that relates to network algorithmics.
%
%In this paper, we will summarize some useful classifications of distributions and stochastic orders to aid
%in the design of randomized algorithms in network algorithmics. 
%In particular, we cover:
%\begin{enumerate}
%\item majorization and Schur convexity, 
%\item negative association,
%\item convex ordering and supermodularity, and
%\item log-concave distributions.
%\end{enumerate}
%We show how these can be applied to various applications in network algorithmics. Most of the 
%results shown here can be found in literature, and so we endeavor to reference the original 
%paper where the result originated from. 
